A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-11-06 summary.tex
adding kz=0 special case in response to inutsuka-san's query
home/office sync

\subsection{Generalization to locally polytropic disks}
We can extend the dusty/adiabatic gas correspondence to 
other fixed equations of state. As an example, consider the locally
polytropic disk 
\begin{align}
  P = K(r,z)\rhog^{\Gamma}, 
\end{align}
where $K$ is a prescribed function and $\Gamma$ is the constant
adiabatic index. Then eliminating $\tepsilon$ from the dust equation
\ref{dusteq} gives 
\begin{align}\label{poly_energy}
  \frac{D P}{D t} &= - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla
  \ln{K} + \mathcal{C},\\
  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right).
\end{align}
Thus the dusty gas behaves like a pure gas with adiabatic index
$\Gamma$. The entropy is given by 
\begin{align}
  \seff = \ln{\left[K^{1/\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
\end{align}
The results of \S\ref{iso_perfect}---\ref{dust_work} remain valid for the strictly 
polytropic disk with $K=$constant, while one sets $c_s^2\to K$ in
\S\ref{dusty_vsi_int}. 
%When $K$ is constant, the corresponding Solberg-Hoiland criteria for
%axisymmetric stability may be obtained from that in standard adiabatic 
%hydrodynamics \citep[e.g.][]{tassoul78} by
%using this definition of entropy in
%Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.

>>>>

\subsection{Generalizations}

\subsubsection{Locally polytropic disks}
We can extend the dusty/adiabatic gas correspondence to 
other fixed equations of state. As an example, consider the locally
polytropic disk 
\begin{align}
  P = K(r,z)\rhog^{\Gamma}, 
\end{align}
where $K$ is a prescribed function and $\Gamma$ is the constant
adiabatic index. Then eliminating $\tepsilon$ from the dust equation
\ref{dusteq} gives 
\begin{align}\label{poly_energy}
  \frac{D P}{D t} &= - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla
  \ln{K} + \mathcal{C},\\
  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right).
\end{align}
Thus the dusty gas behaves like a pure gas with adiabatic index
$\Gamma$. The entropy is given by 
\begin{align}
  \seff = \ln{\left[K^{1/\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
\end{align}
The results of \S\ref{iso_perfect}---\ref{dust_work} remain valid for the strictly 
polytropic disk with $K=$constant, while one sets $c_s^2\to K$ in
\S\ref{dusty_vsi_int}. 
%When $K$ is constant, the corresponding Solberg-Hoiland criteria for
%axisymmetric stability may be obtained from that in standard adiabatic 
%hydrodynamics \citep[e.g.][]{tassoul78} by
%using this definition of entropy in
%Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.

\subsubsection{Higher order corrections as additional stresses}
\begin{align}
\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = -\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\Delta\bm{v}\otimes\Delta\bm{v}\right). 
\end{align}
